---
title: "Lowlands 2022"
author: "Marijn Alta"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
---

```{r Load packages, message=FALSE, echo=FALSE}
library(tidyverse)
library(spotifyr)
library(remotes)
library(usethis)
library(plotly)
library(compmus)
library(DT)
library(forcats)
library(purrr)
library(knitr)
```

```{r Load Playlist, echo=FALSE}
Sys.setenv(SPOTIFY_CLIENT_ID = '5b0d512e353341d1986dd873ae8a7941')
Sys.setenv(SPOTIFY_CLIENT_SECRET = '4306fa250e454c2592ee74cdef497c4e')
ll_playlist <- get_playlist_audio_features("", "5KjYEGVaP7kDz8kR5z4u1d")
```

### (NEW) Tempo curves of the self-similarity matrices of multiple performances of the same piece (Heat Waves)

```{r}
bzt <-
  get_tidy_audio_analysis("1h2AvjqgWC6hakpbGJUb7Z") %>% 
  compmus_align(bars, segments) %>%                     
  select(bars) %>%                                      
  unnest(bars) %>%                                      
  mutate(
    pitches =
      map(segments,
        compmus_summarise, pitches,
        method = "rms", norm = "euclidean"              
      )
  ) %>%
  mutate(
    timbre =
      map(segments,
        compmus_summarise, timbre,
        method = "rms", norm = "euclidean"              
      )
  )
bind_rows(
  bzt %>%
    compmus_self_similarity(pitches, "manhattan") %>%
    mutate(d = d / max(d), type = "Chroma"),
  bzt %>%
    compmus_self_similarity(timbre, "euclidean") %>%
    mutate(d = d / max(d), type = "Timbre")
) %>%
  mutate() %>%
  ggplot(
    aes(
      x = xstart + xduration / 2,
      width = xduration,
      y = ystart + yduration / 2,
      height = yduration,
      fill = d
    )
  ) +
  geom_tile() +
  coord_fixed() +
  facet_wrap(~type) +
  scale_fill_viridis_c(guide = "none") +
  theme_classic() +
  labs(x = "", y = "")

bzt <-
  get_tidy_audio_analysis("1K2RnVhm3eauKnA8PAcPu1") %>% 
  compmus_align(bars, segments) %>%                     
  select(bars) %>%                                      
  unnest(bars) %>%                                      
  mutate(
    pitches =
      map(segments,
        compmus_summarise, pitches,
        method = "rms", norm = "euclidean"              
      )
  ) %>%
  mutate(
    timbre =
      map(segments,
        compmus_summarise, timbre,
        method = "rms", norm = "euclidean"              
      )
  )
bind_rows(
  bzt %>%
    compmus_self_similarity(pitches, "manhattan") %>%
    mutate(d = d / max(d), type = "Chroma"),
  bzt %>%
    compmus_self_similarity(timbre, "euclidean") %>%
    mutate(d = d / max(d), type = "Timbre")
) %>%
  mutate() %>%
  ggplot(
    aes(
      x = xstart + xduration / 2,
      width = xduration,
      y = ystart + yduration / 2,
      height = yduration,
      fill = d
    )
  ) +
  geom_tile() +
  coord_fixed() +
  facet_wrap(~type) +
  scale_fill_viridis_c(guide = "none") +
  theme_classic() +
  labs(x = "", y = "")

bzt <-
  get_tidy_audio_analysis("7GyQCnXcO9d1IQCqx1KOcJ") %>% 
  compmus_align(bars, segments) %>%                     
  select(bars) %>%                                      
  unnest(bars) %>%                                      
  mutate(
    pitches =
      map(segments,
        compmus_summarise, pitches,
        method = "rms", norm = "euclidean"              
      )
  ) %>%
  mutate(
    timbre =
      map(segments,
        compmus_summarise, timbre,
        method = "rms", norm = "euclidean"              
      )
  )
bind_rows(
  bzt %>%
    compmus_self_similarity(pitches, "manhattan") %>%
    mutate(d = d / max(d), type = "Chroma"),
  bzt %>%
    compmus_self_similarity(timbre, "euclidean") %>%
    mutate(d = d / max(d), type = "Timbre")
) %>%
  mutate() %>%
  ggplot(
    aes(
      x = xstart + xduration / 2,
      width = xduration,
      y = ystart + yduration / 2,
      height = yduration,
      fill = d
    )
  ) +
  geom_tile() +
  coord_fixed() +
  facet_wrap(~type) +
  scale_fill_viridis_c(guide = "none") +
  theme_classic() +
  labs(x = "", y = "")
```

***

I chose to look at Heat Waves from Glass Animals. In 2021 Glass Animals released the Heat Waves Expansion Pack which includes the original song, a slow version, a live version, a stripped back version and five remixes of different artists. On this page, the original song, the slow version and the live version are shown.

### (NEW) Histogram of tempi for different groups in your corpus.

***

Histograms for the most popular genres in the corpus.

### (NEW) Tempograms for outliers

***

Tempogram for Dinaledi and ...
    
### Introduction
Do not show in the class. 
Note to self: elaborate on plots and update introduction. 

**What is your corpus, why did you choose it, and what do you think is interesting about it?**

***

A collection of songs of the artists of the lowlands festival in (19xx-?)2022. I chose this topic because I am interested in the development of the lineup of one of my favorite festivals.

**What are the natural groups or comparison points in your corpus (e.g., albums, artists, genres, people)? What do you expect to be different among these groups, what do you expect to be the same, and what are you unsure about?**

- Genres
- Gender of artists
Expect a lot of bands and few female artists.

**How representative are the tracks in your corpus for the groups you want to compare? For example, is there a well-known difference between the live and studio versions of one of the artists, or is Spotify missing an important recording? Or if you are using a personal playlist of yourself or someone else, what percentage of your or their music listening takes place on Spotify? If your corpus is genre-based, how well do you think it covers the genre you are interested in, and what do you think are the most important gaps? There is no such thing as a perfect corpus, and it is important to think from the beginning about its strengths and limitations.**

There might be some artists that are not featured in the playlists, especially in playlists that are older. Niche electronic artists often do not have a Spotify page. Overall I think the playlists are very representative because there are three songs from each of the artists in the playlists.

**Identify several tracks in your corpus that are either extremely typical or atypical. Why do you think that they are so typical or so strange? Several pages of your final portfolio will be dedicated to detailed analyses of specific tracks, so it is important to think about which ones might be most interesting.** 

Typical: popular bands

Atypical: artists with longer tracks or niche genres


### Histogram of keys & keygrams

***

<>
   

### Tempo & energy

```{r}

green <- "#1ed760"

tracks <- 
  ll_playlist %>% mutate(playlist = "Lowlands 2022")

viz1 <- ggplot(tracks, aes(x=tempo, fill=playlist,
                    text = paste(playlist_name)))+
  geom_density(alpha=0.7, color=NA)+
  scale_fill_manual(values=c(green))+
  labs(x="Tempo", y="Density") +
  guides(fill=guide_legend(title="Playlist"))+
  theme_minimal()+
  ggtitle("Distribution of Tempo Data")

ggplotly(viz1, tooltip=c("text"))

```   
 
 ***
 
Test

### Danceability & loudness
```{r}
viz2 <- ggplot(tracks, aes(x=danceability, fill=playlist,
                    text = paste(playlist_name)))+
  geom_density(alpha=0.7, color=NA)+
  scale_fill_manual(values=c(green))+
  labs(x="Danceability", y="Density") +
  guides(fill=guide_legend(title="Playlist"))+
  theme_minimal()+
  ggtitle("Distribution of Danceability Data")

ggplotly(viz2, tooltip=c("text"))
```   

### Chromagram (Dinaledi)
```{r}
track <-
  get_tidy_audio_analysis("0kNOuZYnpu1RiQTWXUUzE8") %>%
  select(segments) %>%
  unnest(segments) %>%
  select(start, duration, pitches)
track %>%
  mutate(pitches = map(pitches, compmus_normalise, "chebyshev")) %>%
  compmus_gather_chroma() %>% 
  ggplot(
    aes(
      x = start + duration / 2,
      width = duration,
      y = pitch_class,
      fill = value
    )
  ) +
  geom_tile() +
  labs(x = "Time (s)", y = NULL, fill = "Magnitude", title = "Dinaledi Chromagram") +
  theme_minimal() +
  scale_fill_viridis_c()
```

***

This song is an outlier in the corpus, because it is the longest song, 8:16. It is a pretty mellow song, with constant energy. The chromogram is pretty evenly distrubeted, which explains the constant energy. 

### Self-similarity (Dinaledi)
```{r}
bzt <-
  get_tidy_audio_analysis("0kNOuZYnpu1RiQTWXUUzE8") %>% 
  compmus_align(bars, segments) %>%                     
  select(bars) %>%                                      
  unnest(bars) %>%                                      
  mutate(
    pitches =
      map(segments,
        compmus_summarise, pitches,
        method = "rms", norm = "euclidean"              
      )
  ) %>%
  mutate(
    timbre =
      map(segments,
        compmus_summarise, timbre,
        method = "rms", norm = "euclidean"              
      )
  )
bind_rows(
  bzt %>%
    compmus_self_similarity(pitches, "manhattan") %>%
    mutate(d = d / max(d), type = "Chroma"),
  bzt %>%
    compmus_self_similarity(timbre, "euclidean") %>%
    mutate(d = d / max(d), type = "Timbre")
) %>%
  mutate() %>%
  ggplot(
    aes(
      x = xstart + xduration / 2,
      width = xduration,
      y = ystart + yduration / 2,
      height = yduration,
      fill = d
    )
  ) +
  geom_tile() +
  coord_fixed() +
  facet_wrap(~type) +
  scale_fill_viridis_c(guide = "none") +
  theme_classic() +
  labs(x = "", y = "")
```

***

Self similarity matrices of the chroma and timbre features of the Dinaledi song. What can be said about the structure? There is a clear 3x3 grid in the chroma plot, which also can be seen less clearly in the timbre plot. What could this mean?

 
### Disussion
 
 ***
 
Conclusions and further work. 
 



